---
title: "Final Project"
author: ""
date: ""
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)                
```


```{r include=FALSE}
library(knitr)
require(mosaic)
#opts_chunk$set(eval=FALSE)
```

```{r load-data}
airline_delay <- read.csv("Airline_Delay_Cause.csv", as.is=T)
```

```{r names}
names(airline_delay)
```
```{r airline-delay}
summary(airline_delay)
```

```{r}
head(airline_delay)
```


###SLR
```{r}
pairs(airline_delay[, c("arr_delay","carrier_delay", "weather_delay", "nas_delay", "security_delay", "late_aircraft_delay")],
      main = "Scatter Plot Matrix of Delay Types",
      pch = 19,  
      col = "blue")  

```
### Linear Regression Model for Arrival delay vs. carrier delay
```{r}
arr_carrier_model <- lm(arr_delay~carrier_delay, data = airline_delay)

summary(arr_carrier_model)
```
```{r}
xyplot(arr_delay~carrier_delay, data = airline_delay, type = c("p","r"))

```

### Linear Regression Model for Arrival delay vs. weather delay
```{r}
arr_weather_model <- lm(arr_delay~weather_delay, data = airline_delay)

summary(arr_weather_model)
```

```{r}
xyplot(arr_delay~weather_delay, data = airline_delay, type = c("p","r"))

```
### Linear Regression Model for Arrival delay vs. national air system (nas) delay
```{r}
arr_nas_model <- lm(arr_delay~nas_delay, data = airline_delay)

summary(arr_nas_model)
```
```{r}
xyplot(arr_delay~nas_delay, data = airline_delay, type = c("p","r"))
```
### Linear Regression Model for Arrival delay vs. security delay
```{r}
arr_sec_model <- lm(arr_delay~security_delay, data = airline_delay)

summary(arr_sec_model)
```

```{r}
xyplot(arr_delay~security_delay, data = airline_delay, type = c("p","r"))
```
### Linear Regression Model for Arrival delay vs. late aircraft delay
```{r}
arr_la_model <- lm(arr_delay~late_aircraft_delay, data = airline_delay)

summary(arr_la_model)
```
```{r}
xyplot(arr_delay~late_aircraft_delay, data = airline_delay, type = c("p","r"))
```
###Late aircraft delay is the main causes of delay

###Plots to check the assumptions of the linear model for arrival delay vs. late aircraft delay

```{r}
xyplot(resid(arr_la_model) ~ fitted(arr_la_model), data=airline_delay, type=c("p","r"))
```
```{r}
histogram(~residuals((arr_la_model)))
```
```{r}
qqmath(~resid(arr_la_model))
```
```{r}
ladd(panel.qqmathline(resid(arr_la_model)))
```
```{r}
library(tidyverse)
# Clean the data by removing rows with NAs
airline_delay <- na.omit(airline_delay)

# Calculate the mean and standard deviation of arrival delays
mean_arr_delay <- mean(airline_delay$arr_delay)
sd_arr_delay <- sd(airline_delay$arr_delay)

# Number of observations
n <- length(airline_delay$arr_delay)

# Calculate the confidence interval
conf_interval <- t.test(airline_delay$arr_delay)$conf.int
conf_interval

# Print the results with more descriptive outputs
```

```{r}
cat("Statistical Summary of Arrival Delays:\n")
cat("Mean Arrival Delay:", round(mean_arr_delay, 2), "minutes\n")
cat("Standard Deviation:", round(sd_arr_delay, 2), "minutes\n")
cat("Number of Observations:", n, "\n")
cat("95% Confidence Interval for the Mean Arrival Delay: [", round(conf_interval[1], 2), ",", round(conf_interval[2], 2), "] minutes\n")
```
```{r}
# Plot histogram of arrival delays with normal curve
hist(airline_delay$arr_delay, breaks = 30, probability = TRUE,
     main = "Histogram of Arrival Delays with Normal Curve",
     xlab = "Arrival Delay (minutes)", col = "lightblue", ylim = c(0, 2.5e-05),
     xlim = c(0, 150000))

# Add normal curve
x <- seq(min(airline_delay$arr_delay), max(airline_delay$arr_delay), length = 100)
y <- dnorm(x, mean = mean_arr_delay, sd = sd_arr_delay)
lines(x, y, col = "red", lwd = 2)

# Add mean line
abline(v = mean_arr_delay, col = "blue", lwd = 2)

# Add confidence interval lines
abline(v = conf_interval[1], col = "green", lwd = 2, lty = 2)
abline(v = conf_interval[2], col = "green", lwd = 2, lty = 2)

# Display the mean and confidence interval on the plot
legend("topright", legend = c(paste("Mean =", round(mean_arr_delay, 2)),
                              paste("95% CI =", round(conf_interval[1], 2), "-", round(conf_interval[2], 2))),
       col = c("blue", "green"), lwd = 2, lty = c(1, 2), bty = "n")
```
```{r}
high_traffic <- filter(airline_delay, arr_flights > 700)
low_traffic <- filter(airline_delay, arr_flights <= 700)
```
```{r}
t_test_result <- t.test(high_traffic$arr_delay, low_traffic$arr_delay, na.rm = TRUE)
t_test_result
```

